---
title: 'World Happiness: A Multivariate Analysis'
author: "Daniel Cooper, Maria Lechleitner, Dahmane Sheikh"
date: "May 2018"
output:
  html_document:
    df_print: paged
---

# Multivariate Normality Assessment: 

The first plot is the QQ plot on raw data, while the second one is when applying power transformation, which will be used for further analysis (factor analysis, pca, cluster analysis, predictions,...)

```{r, echo = TRUE, fig.width=8, fig.height=8}
require(dplyr)

#Using local path as cant' use the general path as this issue constantly appear: 
#cannot open file '../../DATA/full_best_df.csv': No such file or directoryError in file(file, "rt") : cannot open the connection

raw.df <- read.csv("C:/Users/A541U/Desktop/Semester 2/Descriptive R Project/2018_World_Happiness_Multivariate_Analysis/DATA/happiness.csv")

temp_df <- raw.df %>% select(-X,-Country,-Region)
rownames(temp_df) <- raw.df[1:132,2]

n <- nrow(temp_df); p <- ncol(temp_df)

mean <- colMeans(temp_df)
cov <- cov(temp_df)

# ordering mahalanobis distances
d_square <- mahalanobis(temp_df, mean,cov, tol=1e-20)
d_square <- sort(d_square)

# Generating a vector of quantiles
quantiles=(1:n-0.5)/n
x <- qchisq(quantiles,df=p)

# Plot

plot(x, d_square,
     pch=4, xlim=c(0,max(x)*1.2), ylim=c(0,max(d_square)*1.2),
     xlab="Chi-squared quantiles",
     ylab="Mahalanobis distances",
     main="QQ plot - Assessment of multivariate normality")
abline(a = 0, b = 1, col="blue")
text(x,d_square, labels=rownames(temp_df), pos=4,cex=0.7)

rm(cov,d_square,mean,n,p,quantiles,x, temp_df)
```


```{r, echo = TRUE, fig.width=8, fig.height=8}
normalized_df <- read.csv("C:/Users/A541U/Desktop/Semester 2/Descriptive R Project/2018_World_Happiness_Multivariate_Analysis/DATA/full_best_df.csv")

normalized_df <- normalized_df %>% select(-X,-Country,-Region)


#new df to use for further analysis
#full_best_df <- cbind(happiness[,2:3],best_df) 
#write.csv(full_best_df,file="full_best_df.csv")

rownames(normalized_df) <- raw.df[1:132,2]

n <- nrow(normalized_df); p <- ncol(normalized_df)
mean <- colMeans(normalized_df)
cov <- cov(normalized_df)

# ordering mahalanobis distances
d_square <- mahalanobis(normalized_df, mean,cov) # tol=1e-20) is used bc R is detecting singular 
d_square <- sort(d_square)

# Generating a vector of quantiles
quantiles=(1:n-0.5)/n
x <- qchisq(quantiles,df=p)

# Plot

plot(x, d_square,
     pch=4, xlim=c(0,max(x)*1.2), ylim=c(0,max(d_square)*1.2),
     xlab="Chi-squared quantiles",
     ylab="Mahalanobis distances",
     main="QQ plot with Multivariate Normality")
abline(a = 0, b = 1, col="blue")

text(x,d_square, labels=rownames(normalized_df), pos=4,cex=0.7)

rm(cov,d_square, mean,n,p,quantiles,x, raw.df, normalized_df)
```

# Principal Component Analysis:

Firstly, there is a scree plot. Secondly the corrplot. Finally two biplots of variables and countries, respectively.

```{r, echo = TRUE, fig.width=8, fig.height=8}
require(dplyr);require(FactoMineR);require(factoextra)
require(ggcorrplot);require(ggplot2)

# full_best_df <- read.csv(file = "../../DATA/full_best_df.csv")
# this command yields this error: Error in file(file, "rt") : cannot open the connection

full_best_df <- read.csv("C:/Users/A541U/Desktop/Semester 2/Descriptive R Project/2018_World_Happiness_Multivariate_Analysis/DATA/happiness.csv")

#row.names(full_best_df) <- full_best_df$Country

temp_df <- full_best_df %>% 
  select(-X,-Country,-Region,-Happiness_Score,
         -Total_Population,-Mental_and_Substance_Disorder_Index,-Compulsory_Education_in_Years, -Agricultural_Land_Percentage,-Generosity,-Forest_Area_Land_Percentage,
         -Suicide_Index,-Legal_Rights_Index,-Infant_Immunization_Measles_Index, -HIV_Disease_Index, -Alcohol_Consumption_Index)

res.pca <- PCA(temp_df,graph=F) #PCA Function
```

```{r, echo = TRUE, fig.width=8, fig.height=8}
xaxis <- 1:10; eig <- data.frame(res.pca$eig); eig <- eig[1:10,]
temp_for_plot <- data.frame(eig,xaxis)

temp_for_plot$percentage.of.variance <- round(temp_for_plot$percentage.of.variance, digits=2)

ggplot(data=temp_for_plot, aes(x=factor(xaxis), y=percentage.of.variance)) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label= paste0(temp_for_plot$percentage.of.variance, "%")), vjust=-0.3) +
  theme_minimal() + xlab("Dimensions") + ylab("Percentage of explained variances")+
  ggtitle("Scree Plot for PCA")+ theme(plot.title = element_text(hjust = 0.5)) +
  annotate(geom = "text", x = 1, y = 21, label = "Country Development", color = "black",
             angle = 90) +
  annotate(geom = "text", x = 2, y = 20, label = "<---- Air Pollution & Alcohol", color = "black",angle = 90) +
    annotate(geom = "text", x = 3, y = 20, label = "<---- Freedom & Corruption", color = "black",angle = 90)

ggcorrplot(res.pca$var$cos2, lab=T, insig ="blank") + coord_flip() + ggtitle("Quality of Representation per Dimension")

rm(eig,temp_for_plot,xaxis,temp_df)

```

```{r, echo = TRUE, fig.width=8, fig.height=8}
# Color by cos2 values: quality on the factor map
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE) # Avoid text overlapping

row.names(res.pca$ind$coord) <- full_best_df$Country

#still need to solve the layout error for this graph in this rmd. 
fviz_pca_ind(res.pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) # Avoid text overlapping (slow if many points)

dimdesc(res.pca)

```
 
# Linear Regression (Single and Multiple) based on PCA:

Simple Linear Regression using only the 1st Principal Component:
```{r, echo = TRUE, fig.width=8, fig.height=8}
pc1lm <- lm(full_best_df$Happiness_Score ~ res.pca$ind$coord[,1])
summary(pc1lm) #R_Square of 65% with just PC1
plot(res.pca$ind$coord[,1],full_best_df$Happiness_Score)
abline(pc1lm)
#text(res.pca$ind$coord[,1],full_best_df$Happiness_Score, label=full_best_df$Country)

temp_plot <- data.frame(pc1 = res.pca$ind$coord[,1], Happiness_Score = full_best_df$Happiness_Score)

ggplot(temp_plot, aes(x=pc1, y=Happiness_Score)) + geom_point() + geom_smooth(method="lm", col="red") +
  ggtitle(label="Simple Linear Regression",
  subtitle = labs(title =paste("Adj R2 =",signif(summary(pc1lm)$adj.r.squared, 5),
                               "Intercept =",signif(pc1lm$coef[[1]],5 ),
                               " Slope =",signif(pc1lm$coef[[2]], 5),
                               " P =",signif(summary(pc1lm)$coef[2,4], 5)))) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust =0.5)) +
  xlab("1st Principal Component (Country Development)") + ylab("Happiness_Score")
```

Multiple Linear Regression with the first 3PCs:
```{r, echo = TRUE, fig.width=8, fig.height=8}
predictors3pcs <- data.frame(res.pca$ind$coord[,1:3],happiness_score = full_best_df$Happiness_Score)
fit3pcs <- lm(happiness_score~ Dim.1 + Dim.2 + Dim.3, data=predictors3pcs)
summary(fit3pcs) #R_Square of 78%
```


# Linear Regression (Single and Multiple) based on FA:

Simple Linear Regression using only the 1st Factor:

```{r, echo = TRUE, fig.width=8, fig.height=8}
#Get data
library(psych)
FA_Happiness <- principal(selected, nfactors=3, rotate="varimax")
FA1_Scores <- FA_Happiness$scores[,1]
HP_Scores <- full_best_df$Happiness_Score

#Regression
FA1lm <- lm(HP_Scores ~ FA1_Scores) #linear model
summary(FA1lm) #R_Square of 31% with just PC1
plot(FA1_Scores, HP_Scores)
abline(FA1lm)
#text(res.pca$ind$coord[,1],full_best_df$Happiness_Score, label=full_best_df$Country)

temp_plot <- data.frame(Factor1 = FA1_Scores, Happiness_Score = full_best_df$Happiness_Score)

ggplot(temp_plot, aes(x=Factor1, y=Happiness_Score)) + geom_point() + geom_smooth(method="lm", col="red") +
  ggtitle(label="Simple Linear Regression",
  subtitle = labs(title =paste("Adj R2 =",signif(summary(FA1lm)$adj.r.squared, 5),
                               "Intercept =",signif(FA1lm$coef[[1]],5 ),
                               " Slope =",signif(FA1lm$coef[[2]], 5),
                               " P =",signif(summary(FA1lm)$coef[2,4], 5)))) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust =0.5)) +
  xlab("1st Factor (Country Development)") + ylab("Happiness Score")

```


Simple Linear Regression using 3 Factors:

```{r, echo = TRUE, fig.width=8, fig.height=8}
#Get data
FA_Scores <- FA_Happiness$scores
dfFAScores <- data.frame(FA_Scores)
sumScores <- dfFAScores$RC1+dfFAScores$RC2+dfFAScores$RC3
HP_Scores <- full_best_df$Happiness_Score

#Regression
predictors3fas <- data.frame(FA_Scores,HP_Scores)
fit3fas <- lm(HP_Scores~ RC1 + RC2 + RC3, data=predictors3fas)
summary(fit3fas) #R_Square of 78%

#Plots
plot(sumScores, HP_Scores)
abline(fit3fas)

temp_plot <- data.frame(Factors = sumScores, Happiness_Score = full_best_df$Happiness_Score)

#not correct yet!
ggplot(temp_plot, aes(x=Factors, y=Happiness_Score)) + geom_point() + geom_smooth(method="lm", col="red") +
  ggtitle(label="Simple Linear Regression",
  subtitle = labs(title =paste("Adj R2 =",signif(summary(fit3fas)$adj.r.squared, 5),
                               "Intercept =",signif(fit3fas$coef[[1]],5 ),
                               " RC1 =",signif(fit3fas$coef[[2]], 5),
                               " RC2 =",signif(fit3fas$coef[[3]], 5),
                               " RC3 =",signif(fit3fas$coef[[4]], 5),
                               " P =",signif(summary(fit3fas)$coef[4,4], 5)))) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust =0.5)) +
  xlab("Factors") + ylab("Happiness Score")
```
```